### 2024-01-24 manuscript.tex
Update manuscript.tex

Female Genital Cutting Is a Social Coordination Norm in Kenya, Mali, Nigeria and Sierra Leon
(or:  Cutting or Not:  Depending on what others will undergo the practice than what they think about the practice)
\cfcmt{This title is a response to Efferson’s title}

@@

Female genital cutting (FGC), circumcision or mutilation, is a form of female's genital mutilation cite{} and mostly viewed as violation of human rights \cite{WHO97, WHO08} \cfonly{add and update}.  The commonly international position of FGC is about abolishment than intervention of the practice \cite{KhosBane17, Mack96, Toub94, UNIC16, WHO97}\cfonly{update}.  It is estimated, between 20xx to 20xx, that more than xxx million women and girls have undergone female genital cutting (FGC), mostly in Africa and the Middle East, and among them, xx million are girls below age 15 \cfonly{based on a UNICEF global database \cite{AdnrLesc16, UNIC16} \cfcmt{I think we can delete these: including Demographic and Health Surveys (DHS), Multiple Indicator Cluster Surveys (MICS), and other nationally representative surveys} adding KandEzej18? and update}.  The UNICEF global database predicted that another xx million girls and women were at risk by 20xx  \cite{AdnrLesc16, UNIC16, update}.  Progress of intervention is impeded because of the complicated FGC history \cite{BergDeni12, Cami16} and of not keeping up with the population growth \cite{KhosBane17, UNIC16}, and no clear change was concluded \cite{KandShel19} \cfonly{update}.  Following Tostan \cite{Tost15}, we used cutting instead of circumcision without diminishing or excusing the negative impact of the practice while recognizing a sense of self-awareness in making decision about FGC \cite{KhahBark09, JohnEsse10, Meye00, PariSaru18, Shel01, Tost15}. \cfonly{less cites}

FGC is embodied with multiple approaches from cultural and religious identity in one end and public health and human rights concerns on the other. \cite{AhmeKare18, BergDeni13a, Boyl02, BoylMcMo02, Grue01, KhosBane17, KimaShell18, McCh15, SchuLien13, WHO12}. \cfonly{too many cites}.  Knowledge and attitudes in supporting FGC did not necessarily translate to support of abandoning the practice nor the intention not to cut daughters \cite{ChegAske04}.  Understanding reasons and values people attached to FGC is essential to comprehend the persistence of the practice \cite{Bicc10, Cami15, CislHeis18, Mack96}.  In this study, we focused on social values of FGC to understand the practice in communities.  There is a considerable interest in applying social norm, a form of values driving behavioural patterns which are self-enforcing within a group \cite{Youn15} to the intervention of public health behaviours \cite{Ajze91, Ajze02, CislHeis18a, MillPrin16, MollRima10, RimaLapi15} \cfonly{focus on FGC instead of the health behaviours in general}, and elaborating social norms of FGC provides a crucial venue in negotiating sustainable changes of the practice \cite{Bicc10, BiccMari15, BoylCorl010, DuncWand11, Drol11, EffeVogt15, Hayf05, HayfTrin11, Grue05, Hodg11,KandNwak09, Mack96, Mack00, MackLeJe08, OdukAfol17, RimaLapi15, Youn02, Youn11, UNIC10, UNIC13} \cfonly{less}.

FGC norms are deeply internalized and morally embraced \cite{SchuLien13}.  Social convention theory, a leading theory probing FGC tradition (e.g., see \cite{BoylMcMo02, BoylCorl10, FreyJohn07, FrieMahm13, Hayf05, KandMwek09, Mack96, Mack06, ReigGonz14, YirgKass12}) initially proposed that FGC was a marriageability convention through coordination \cite{Mack96} and extending to social norms, such as religious obligation, adolescent rite of passage, and female honour and modesty to control marriage fidelity and social prestige \cite{Mack00, MackLeJe08, update}.   The social convention theory regards FGC a social behaviour interdependent among and coordinated by members in their social network \cite{Mack00, MackLeJe08, ShelWand11}.  As no coordination game was observed in marital expectation in community with FGC tradition \cite{EffeVogt15}, studies showed that factors other than marriageability, such as \cfonly{examples}, have underlined women’s  decisions of FGC \cite{AlcaGonz13, BellNova15, EffeVogt15, Hayf05, Mack09, PashPonn16, Rima08, ShelWand11, more?}.  FGC was also viewed as a form of social capital, a value respected by women for inclusion within their social groups \cite{ShelWand11}.   The practice was strongly correlated with social expectation (i.e., what other people do or what other people believe one should do)\cite{BiccMari15}.   Basically, when FGC norms within the communities are strong, individuals tend to self-enforce and coordinate such norms \cite{Ajze02, Hayf05, KandNwak09, KandShel19, Mack96, Mack06, MackLeJe08, ThomMadd92} \cfonly{add \cite{Bicc15}}.

Although FGC norms were studied as aforementioned, there was limited analysis of how individual normative attitudes (e.g., attitudinal position of FGC values) and behavioural intention intertwining with the group norm.  Such information is expected to extend understandings of the persistence of FGC practice.  We believe this is the first FGC study using modules of FGC normative attitudes to model the intention by utilizing FGC norm at a community level.

@@

The research presumption was based on the studies aforementioned that women’s intention on practicing FGC on their daughters was dependent on other women’s intention of this practice \cfonly{descriptive norms?!}. We used index of attitudinal evaluation of FGC benefits to represent FGC norm.  Attitudes is a sensible proxy of social norm \cfonly {norm theory \cite{Ajze91, Ajze02, Bicc10, BiccMari15, Mack96, Mack00, MackLeJe08, RimaLapi15, Youn11} and has been applied on fgc studies \cite{CislHeis18, EffeVogt15, Harf06, ModrLiu13, PashPonn16, ShelHern06,}} \cfonly{check cites}.  However, there were little quantitative research on the relationship between attitudes and intention of circumcising daughters at both an individual and community level while taking beliefs of FGC benefits into account.

Three models were constructed in response to the three research questions mentioned above:  the daughter model, the continuance model and the persistence model.  The main one is the “daughter” model, which analyzed associations of beliefs of FGC benefits (see the list of FGC benefits at table xx) and behavioral intention to have their daughters receiving FGC.  Additionally, we decomposed the model into two “structural” models:  structural one “continuance” model to study women’s beliefs of FGC benefits and FGC continuance; and structural two “persistence”  model \cf{or “mixed” model?} to examine the association of beliefs of FGC benefits, FGC continuance on intention of genitally cutting daughters.  We used the two structural models to further understand how such positions change the main model.

@@

We analyzed the secondary data of women aged 15-49 from the Demographic and Health Surveys (DHS).  The data was selected when it met the research criteria:  countries with high FGC prevalence \cite{UNIC16}, and with the modules of FGC status and benefit and gender awareness. Four countries qualified under this criteria with the most recent data sets: Kenya 2008/9, Mali 2006, Nigeria 2008 and Sierra Leone 2008 \cfcmt{checked and updated in Nov 6, 2023.}   Although the datasets seemed outdated, but it should not affect the study object which was about FGC norms and behaviors.  \cfcmt{See https://github.com/fishforwish/fgc/blob/master/README.md for updated info}.

Only women with daughters for for FGC options were included in the the daughter model and the persistence model.   The continuance model included all the women in the samples; and that resulted in xxx, xxx, xxx and xxx, and 7861, 13071, 18311 and 7231 in Kenya, Mali, Nigeria and Sierra Leon respectively.  \cfcmt{FGC module at \url{https://dhsprogram.com/pubs/pdf/DHSQMP/DHS5_Module_Female_Genital_Cutting.pdf}}

@@

The main response variables were woman's reported intention to modify their daughters’ genitals in the daughter model and the persistence model, and woman's attitude on FGC future in the continuance model.  The main predictors were woman's FGC status and beliefs of FGC benefits (see supp for the benefit questionnaire).  The beliefs of FGC benefits were quantified using average score \cfcmt{pls confirm} to identify the strength of fgc beliefs associating with FGC practice.

Co-variables in the models were basic social demographic factors (i.e., country, education, job, residence (rural vs. urban), religion, marital status, age, media use, and wealth) in addition to gender awareness.  Gender awareness (see — for questions of gender awareness proxy) was included to test if women’s gender awareness interacted with their intention of cutting daughters’ genitals and their position of FGC practice overall \cfonly{cite{Dell04, FrieMahm13,KhosBane17, Lewi95, Lewi09, Meye00, Morr08, Njam04, UNIC16, YirgKass12, Youn02} check and cut}.  Media use and gender awareness were scored.  Cluster ID (i.e., villages where the samples were from) and ethnicity (see the list of ethnicity recode at table xx in appendix) \cfonly{with a footnote on how we recoded it} were treated as random variables.  Ethnicity was applied as a random effect based on previous study which showed little differences in results using individual rather than cluster ethnicity \cite{Hayf05}.

Regarding potential community impact on the practice of FGC, mothers' intention on having FGC on daughters, FGC continuance, and FGC beliefs. The following also tested at a community level:  education, wealth, media use, FGC beliefs, gender awareness and FGC prevalence at the community level \cfcmt{on a cluster level, not national, right?} with cluster as a proxy to represent a community level of impact \cite{AligRen06, BoylSvec19, Hayf05, Krav02}.

\cfcmt{Regarding FGC benefits modules, there were 9 questions.  Kenya had all the 9, Mali and SL 7 missing promiscuity and STD, NG 8 missing STD).  Should we drop STD since 3 out of 4 missed this variable?}
\jdcmt{Ideally, we would make ethnicity a random effect, but we are back to the Gilmour problem I guess.} \cfcmt{Ethnicity is an important factor (see Hayf05), more so than religion, associating with FGC status, and I don't think it shall be coded as a random factor.  But as J said, it is too hard! } \cfcmt{Whatever we decide, we need a reason in this section to justify our decision. After all, ethnicity is better to be a fixed factor but we can’t due to it breaking down the models!}
\cfcmt{“explicitly including ethnicity is a means of assessing another dimension of the collective aspect of circumcision behavior” \cite{Hayf05}}
\cfcmt{Bayesian model \cite{KandNwak09} “Conversely, one cannot assume that the clusters selected in each district are fully representative of the states in which they are located because surveys only attempted to generate a fully representative sample at the regional level. Consequently, the spatial analysis will be affected by some random fluctuations.  Some of this random variation can be reduced through structured spatial effects because it includes neighboring observations in the analysis. However, it should be pointed out that such a spatial analysis should preferably be applied to census data, where the precision of the spatial analysis would be much higher." (p. 788)}

@@

\cfonly{Attitude is “personal evaluation of a behaviour” \cite{Ajze91}}

@@

We used cumulative link mixed models (CLMMs) in the statistics package R \cite{Rstats,Rpackage_ordinal} to analyze the models.  The CLMM framework allows us to model a binary or ordinal response variable (i.e., intention of applying fgc to daughters and whether to continue FGC practice), while treating clusters and ethnicity as random effects.  
We subtracted respondent (-1) from the cluster when testing the community effect \cfcmt{Please rewrite this. Also, do we need to mention how we treat cluster with only one sample (if that happens).}
\cfcmt{Please explain why we didn’t do sampling weight.}
\cfcmt{Do we explain conveyance of uncertainty}

@@

@@

>>>>

Female Genital Cutting Is a Social Coordination Norm in Kenya, Mali, Nigeria and Sierra Leon
\cfcmt{This title is a response to Efferson’s title}

@@

Female genital cutting (FGC), circumcision or mutilation is a form of female's genital mutilation \cfonly{cite}, Following Tostan \cite{Tost15} \cfonly{update to make sure Tostan still perfers the term}, we used cutting instead of circumcision in this study without excusing the negative impact of the practice while recognizing a sense of self-awareness in making decision about FGC \cite{KhahBark09, JohnEsse10, Meye00, PariSaru18, Shel01, Tost15}. \cfonly{less cites and update cites}.  It is estimated that more than xxx million women and girls will have undergone female genital cutting (FGC), mostly in Africa and the Middle East, between 20xx to 20xx; and among them, xx million are girls below age 15 \cfonly{based on a UNICEF global database \cite{AdnrLesc16, UNIC16} \cfcmt{check cites and adding KandEzej18? update}. The prevalence might be underreported due to a deeply embedded cultural practice that is mostly illegal \cite{StevKell22}.  FGC is embodied with multiple approaches from cultural and religious identity in one end, and public health and human rights concerns on the other \cite{AhmeKare18, BergDeni13a, Boyl02, BoylMcMo02, Grue01, KhosBane17, KimaShell18, McCh15, SchuLien13, WHO12} \cfonly{too many cites and update}.  The commonly international position of FGC is about abolishment than intervention of the practice \cite{KhosBane17, Mack96, Toub94, UNIC16, WHO97}\cfonly{update}.  Progress of intervention is impeded because of the complicated FGC history \cite{BergDeni12, Cami16} and of not keeping up with the population growth \cite{KhosBane17, UNIC16}.

Knowledge and attitudes in supporting FGC abolishment did not necessarily translate to abandon the practice nor the intention not to modify women's genitals \cite{ChegAske04, Bicc10, Cami15, CislHeis18, Mack96} \cfonly{more cites}.  Understanding reasons and values attached to FGC is essential for comprehending the persistence of the practice \cite{Bicc10, Cami15, CislHeis18, Mack96, Youn15} \cfonly{update}.  There is a considerable interest in applying social norm, a form of values driving behavioural patterns which are self-enforcing within a group \cite{Youn15}, to the intervention of public health behaviours \cite{Ajze91, Ajze02, CislHeis18a, MillPrin16, MollRima10, RimaLapi15} \cfonly{focus on FGC instead of the health behaviours in general, and update cites}.  FGC norms are deeply internalized and morally embraced \cite{SchuLien13} \cfonly{update}.  Social convention theory, a leading theory probing FGC tradition (e.g., see \cite{BoylMcMo02, BoylCorl10, FreyJohn07, FrieMahm13, Hayf05, KandMwek09, Mack96, Mack06, ReigGonz14, YirgKass12} \cfonly{update}) initially proposed that FGC was a marriageability convention through coordination \cite{Mack96} and extended to broader social norms, such as social prestige, adolescent rite of passage, and female honour and modesty \cite{Mack00, MackLeJe08} \cfonly{more cites}.   The social convention theory regards FGC a social behaviour interdependent among and coordinated by members in their social network \cite{Mack00, MackLeJe08, ShelWand11} \cfonly{more cites}.  As no coordination game was observed in marital expectation in community with FGC tradition \cite{EffeVogt15}\cfonly{more cites}, studies showed that factors other than marriageability, such as \cfonly{examples}, have underlined women’s  decisions of FGC \cite{AlcaGonz13, BellNova15, EffeVogt15, Hayf05, Mack09, PashPonn16, Rima08, ShelWand11, more?}.  FGC was also viewed as a form of social capital, a value respected by women for inclusion within their social groups \cite{ShelWand11}.   The practice was strongly correlated with social expectation (i.e., what other people do or what other people believe one should do)\cite{BiccMari15} \cfonly{more cites}.   Basically, when FGC norms within the communities are strong, individuals tend to self-enforce and coordinate such norms \cite{Ajze02, Hayf05, KandNwak09, KandShel19, Mack96, Mack06, MackLeJe08, ThomMadd92} \cfonly{add \cite{Bicc15}}\cfonly{more cites}.

Understanding social norms of FGC provides a venue in negotiating sustainable changes of the practice \cite{Bicc10, BiccMari15, BoylCorl010, DuncWand11, Drol11, EffeVogt15, Hayf05, HayfTrin11, Grue05, Hodg11,KandNwak09, Mack96, Mack00, MackLeJe08, OdukAfol17, RimaLapi15, Youn02, Youn11, UNIC10, UNIC13} \cfonly{less and update}.  In this study, we anlayzed types of social values of FGC perceived in communities and the associations of social values and FGC practice.  Although FGC norms were studied as aforementioned, there was limited analysis of how individual normative attitudes (e.g., attitudinal position of FGC values) and behavioural intention intertwining with the group norm.  Such information is expected to extend understandings of the persistence of FGC practice.  We believe this is the first FGC study using modules of FGC normative attitudes to model the FGC intention by utilizing the norm in the context of community influence.

@@

There are two research questions:  what is the association of daughter's FGC future and mothers' FGC beliefs, and what is the association of FGC continuance and FGC beliefs. The research presumption was based on the studies aforementioned that women’s FGC intention on their daughters was dependent on other women’s attitudes on this practice \cfonly{descriptive norms?!}.  Attitude is a personal evaluation of behaviour \cite{Ajze91}, a sensible proxy of social norm \cfonly {norm theory \cite{Ajze91, Ajze02, Bicc10, BiccMari15, Mack96, Mack00, MackLeJe08, RimaLapi15, Youn11}\cfonly{cut cites}}.  It is an important index used to evaluate fgc practices \cite{CislHeis18, EffeVogt15, Harf06, ModrLiu13, PashPonn16, ShelHern06,}} \cfonly{check cites}.  However, there was little quantitative research on the relationship between attitudes and intention of circumcising daughters at both an individual and community level while taking beliefs of FGC benefits into account.

Two main models were constructed in response to the two research questions mentioned above:  the daughter model, and the continuance model.  In addition, we also tested a “persistence” model, to examine the association of FGC benefits and FGC continuance on mothers' intention on FGC their  daughters.  Both of the continuance model and the persistence model were purposed to further understand how the position of the main model could be changed under the two structural models.

@@

We analyzed the secondary data of women aged 15-49 from the Demographic and Health Surveys (DHS) when it met the research criteria:  countries with high FGC prevalence \cite{UNIC16}, and with the modules of FGC status, FGC benefit, and gender awareness. As a result, four countries of the most recent datasets were selected: Kenya 2008/9, Mali 2006, Nigeria 2008 and Sierra Leone 2008 \cfcmt{checked and updated in Nov 6, 2023.}   Although the datasets seemed outdated, it should not affect the study object to understand FGC norms and behaviors.  \cfcmt{See https://github.com/fishforwish/fgc/blob/master/README.md for updated info}.

Only women with daughters for FGC options were included in the the daughter model and the persistence model.   The continuance model included all the women in the datasets; and that resulted in xxx, xxx, xxx and xxx, and 7861, 13071, 18311 and 7231 in Kenya, Mali, Nigeria and Sierra Leon respectively.  \cfcmt{FGC module at \url{https://dhsprogram.com/pubs/pdf/DHSQMP/DHS5_Module_Female_Genital_Cutting.pdf}}

@@

The main responses and predictors in the models are as: FGC benefits (predictor) -->  to FGC daughters (response) in the daughter model, FGC benefits --> FGC future in the continuance model, and FGC benefits and FGC continuance --> to FGC daughters in the persistence model.  The beliefs of FGC benefits were quantified using average score \cfcmt{pls confirm} to identify the strength of FGC beliefs associating with FGC practice.

The co-variables were basic social demographic factors (i.e., country, education, job, residence (rural vs. urban), religion, marital status, age, media use, and wealth) in addition to gender awareness.  Gender awareness (see — for questions of gender awareness proxy) was included to test if women’s gender awareness interacted with their intention of FGC daughters and their position of FGC practice overall \cfonly{cite{Dell04, FrieMahm13,KhosBane17, Lewi95, Lewi09, Meye00, Morr08, Njam04, UNIC16, YirgKass12, Youn02} check and cut}.  Media use and gender awareness were also scored.  Cluster ID (i.e., villages where the samples were from) and ethnicity (see the list of ethnicity recode at table xx in appendix) \cfonly{with a footnote on how we recoded it} were treated as random variables.  Ethnicity was treated as a random effect based on previous study which showed little differences in results using individual rather than cluster ethnicity \cite{Hayf05}.

Regarding potential community impact on the practice of FGC, the following factors were also tested at a community level:  mothers' intention on FGC daughters, FGC continuance, and FGC benefits as well as the basic social-demographic factors:  education, wealth, media use, gender awareness and FGC prevalence \cfcmt{to confirm:  it's on a cluster level, not national.  Correct?} with cluster as a proxy to represent a community level of impact \cite{AligRen06, BoylSvec19, Hayf05, Krav02}.

\cfcmt{Regarding FGC benefits modules, there were 9 questions.  Kenya had all the 9, Mali and SL 7 missing promiscuity and STD, NG 8 missing STD).  Should we drop STD since 3 out of 4 missed this variable?}
\jdcmt{Ideally, we would make ethnicity a random effect, but we are back to the Gilmour problem I guess.} \cfcmt{Ethnicity is an important factor (see Hayf05), more so than religion, associating with FGC status, and I don't think it shall be coded as a random factor.  But as J said, it is too hard! } \cfcmt{Whatever we decide, we need a reason in this section to justify our decision. After all, ethnicity is better to be a fixed factor but we can’t because it breaks down the models!}

\cfcmt{“explicitly including ethnicity is a means of assessing another dimension of the collective aspect of circumcision behavior” \cite{Hayf05}}
\cfcmt{Bayesian model \cite{KandNwak09} “Conversely, one cannot assume that the clusters selected in each district are fully representative of the states in which they are located because surveys only attempted to generate a fully representative sample at the regional level. Consequently, the spatial analysis will be affected by some random fluctuations.  Some of this random variation can be reduced through structured spatial effects because it includes neighboring observations in the analysis. However, it should be pointed out that such a spatial analysis should preferably be applied to census data, where the precision of the spatial analysis would be much higher." (p. 788)}

@@

@@

We used cumulative link mixed models (CLMMs) in the statistics package R \cite{Rstats,Rpackage_ordinal} to analyze the models.  The CLMM framework allows us to model a binary or ordinal response variable (i.e., intention of applying FGC to daughters and whether to continue FGC practice), while treating clusters and ethnicity as random effects.  
We subtracted respondent (-1) from the cluster when testing the community effect \cfcmt{Please rewrite this. Also, do we need to mention how we treat cluster with only one sample (if that happens).}
\cfcmt{Please explain why we didn’t do sampling weight.}
\cfcmt{Do we explain conveyance of uncertainty}

@@

Norms can be rooted in what one thinks, passively or actively, others think which can be based on one's action or the communication.

@@

Suggested reviewers:
* Kerrie Stevenson (ku.ca.mthsl@nosnevets.eirrek): Faculty of Public Health and Policy, London School of Hygiene and Tropical Medicine, London, United Kingdom; Institute of Health Informatics, University College London, London, United Kingdom
* Brenda Kelly, Oxford University Hospitals NHS Foundation Trust, The Oxford Rose Clinic, Oxford, United Kingdom
